---
output: github_document
---

# zonebuilder

A rust crate for building zones.

It is an experimental and in-progress project to implement the functionality in the [`zonebuilder`](https://zonebuilders.github.io/zonebuilder/) R package in the systems programming language Rust.

Why?

- It should eventually enable more people to benefit from free and open source software for creating zoning systems because Rust enables the creation of binaries for Windows, Mac and the free and open source Linux operating system on which the package was originally developed (Rust can also compile to [WASM](https://webassembly.org/) enabling complex applications such as [A/B Street](https://github.com/a-b-street/abstreet) to run in browser --- the thinking being if that can run in browser surely as simple application to build zones can!)
- Computational efficiency: the process of building zones is not particularly computationally intensive but this Rust crate may eventually be fast and quick to install and use, possibly from higher level languages such as R using Rust interfaces such as [`extendr`](https://github.com/extendr/extendr)
- For fun and education: as a simple crate it serves as a good way to show how Rust code is organised and how it works

To reproduce the example shown here you need to have the rust toolchain installed.

Assuming you do, you can run the code as follows

### Clone the repo

```{r, echo=FALSE}
unlink("zonebuilder-rust", recursive = TRUE, force = TRUE)
```


```{r, engine='bash'}
git clone https://github.com/zonebuilders/zonebuilder-rust.git
cd zonebuilder-rust
git checkout circles
```


### Run the CLI

The `zonebuilder` software is shipped as a binary command line interface that can by called from Windows, Mac and Linux system shells as follows:

```{r, engine='bash'}
cargo build
```

You can see instructions on using the tool with the following command:

```{r, engine='bash'}
./target/debug/zonebuilder -h
```

Let's try making zones with fewer segments:

```{r, engine='bash'}
./target/debug/zonebuilder -s 3 > zones.geojson
```


The result:

```{r, echo=FALSE}
z = sf::read_sf("zones.geojson")
plot(z)
```


```{r, engine='bash'}
./target/debug/zonebuilder --precision 0 > zones.geojson
```

Results in this:

```{r, echo=FALSE}
z = sf::read_sf("zones.geojson")
plot(z)
```






```{r, engine='bash'}
cargo run > zones.geojson
```

Take a look at the output:

```{r, engine='bash'}
head -n 20 zones.geojson
```

Then read in the GeoJSON file with another tool, e.g. R (this step runs from an R console that has the `sf` library installed):

```{r circle}
zones = sf::read_sf("zones.geojson")
plot(zones)
# interactive version:
# mapview::mapview(zones)
file.remove("zones.geojson")
```

You can generate the same plot in R with the `zonebuilder` package as follows:

```{r rversion}
zones = zonebuilder::zb_zone(x = "london", n_circles = 5)
plot(zones$geometry)
```


<!-- ## Tidy up -->

```{r engine='bash', echo=FALSE, eval=FALSE}
cd 
pwd
# rm -rvf zonebuilder-rust
```



<!--
The crate template was made with the following command:

```bash
cargo new --lib zonebuilder
```

```bash
mv -v zonebuilder/* .               
# renamed 'zonebuilder/Cargo.toml' -> './Cargo.toml'
# renamed 'zonebuilder/src' -> './src'
```

Edit the .rs files in src folder.

Then run:

```bash
cargo test
```

-->
